---
title: "Tema_4_supuestos"
author: "Raymond L. Tremblay"
date: "2/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





Add and activate packages.
NOTE THE NEW FUNCTION to install all packages ONLY IF Needed.....

```{r packages, include=FALSE}
#----Install and Load Packages-----
#install.packages("car", dependencies = TRUE)
#install.packages("ggplot2")
#install.packages("pastecs")
#install.packages("psych")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(car, pastecs, psych, ggplot2)


library(car) # Companion to applied regression
library(ggplot2)  # Data Visualization
library(pastecs) # Space time Ecological Series
library(psych)   # Procedures for Psychological, psychometric and Personality Research
#library(modeest) # package to calculate the mode of a dataset
```


Visualize the distribution of the data "Histogram"


Los datos son de un festival de Rock y Heavy Metal en UK
La categoria son
0 = You smell like a rotting corpse
4 = You smell like of sweet roses


```{r histogram}

library(readr)
DownloadFestival <- read_csv("Data/DownloadFestival.csv")
```

```{r}
dlf <- DownloadFestival

dlf=subset(DownloadFestival, day1<5)
head(dlf)
tail(dlf)

hist.day1 <- ggplot(dlf, aes(day1)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") 

hist.day1+ 
  labs(x="Hygiene score on day 1", y = "Density")

hist.day1

ggsave("histogram_festival_hygiene.pdf") # Can be either be a device function (e.g. png()), or one of "eps", "ps", "tex" (pictex), "pdf", "jpeg", "tiff", "png", "bmp", "svg" or "wmf" (windows only)

```

#Remove the outlier from the day1 hygiene score
dlf$day1 <- ifelse(dlf$day1 > 5, NA, dlf$day1)
df$Column = ifelse(df$column_to_be evaluated, replace_with_NA, otherwise_leave_as_before)


```{r remove outlier}
dlf$day1 <- ifelse(dlf$day1 > 5, NA, dlf$day1)


# Note here that we use ..density..  # CUAL ES LA DIFERENCIA entre densidad y frecuencia?  

hist.day1 <- ggplot(dlf, aes(day1)) + 
  theme(legend.position = "none") + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") + 
  labs(x="Hygiene score on day 1", y = "Density")

hist.day1

```



The Normal Distribution

https://en.wikipedia.org/wiki/Normal_distribution


1. visualization if the observed distribution follows a theoretical normal distribution
2. Test to determine if the observed distribution follows a theoretical distribution


 $$P(x)=\frac{1}{{\sigma\sqrt{ 2\pi}}}{e}^{\frac{{(x-µ)}^{2}}{{2\sigma}^{2}}}$$
 




```{r adding the normal curve}
# Ahora añadir la linea de distribución normal

hist.day1 + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(dlf$day1,na.rm = TRUE), 
                sd = sd(dlf$day1 , na.rm = TRUE)), 
                colour = "red", size = 1)

shapiro.test(dlf$day1)  # don't use with more than 40 a 200 data point
length(dlf$day1)
```

A third method

3. The QQ plot method

```{r qqplot}
#Q-Q plot for day 1:
qqplot.day1 <- qplot(sample = dlf$day3, stat="qq")
qqplot.day1

```



Add a straight line on the qqplot
```{r qqplot and line}

# This function is to add a straight line through the qqplot
qqplot.data <- function (vec) # argument: vector of numbers
{
  # following four lines from base R's qqline()
  y <- quantile(vec[!is.na(vec)], c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  int <- y[1L] - slope * x[1L]
  
  d <- data.frame(resids = vec)
  
  ggplot(d, aes(sample = resids)) + 
    stat_qq() + 
    geom_abline(slope = slope, intercept = int, color="red")
  
}

qqplot.data(dlf$day3)


ggsave("qqplot.png")
```

```{r quantifying normality}

dlf=DownloadFestival
#Quantifying normality with numbers
library(psych)		#load the "psych" library, if you haven't already, for the describe() function.

#Using the describe() function for a single variable.
psych::describe(dlf$day2)

```
Que es la varianza?
The variance
$$s^{ 2 }=\frac { \sum _{ i=1 }^{ n }{ (x_{ i }-\bar { x } ) } ^{ 2 } }{ n-1 } $$

Que es la desviación estandard

The standard deviation of the mean
$$s=\sqrt { s^{ 2 } } $$








Multiple variables

```{r multiple variables}
#Two alternative ways to describe multiple variables.
psych::describe(cbind(dlf$day1, dlf$day2, dlf$day3))
psych::describe(dlf[,c("day1", "day2", "day3")])

```


Dicover the Mode. La moda en R

```{r mode}
# the mode
library(modeest)

mfv(dlf$day1, method="mfv")
mfv(dlf$day2, method="mfv")
mfv(dlf$day3, method="mfv")



```




Test of normality, Shapiro Wilks Test

```{r test of normality}

library(pastecs)
stat.desc(dlf$day3, basic = FALSE, norm = TRUE) # "norm=TRUE"" is to calculate the Shapiro Wilk Test
```

```{r description by variable}
stat.desc(cbind(dlf$day1, dlf$day2, dlf$day3), basic = FALSE, norm = TRUE)
```

```{r rounding}
round(stat.desc(dlf[, c("day1", "day2", "day3")], basic = FALSE, norm = TRUE), digits = 3)

```

# Levene's Test: testing for assumption of normality among groups

```{r levene test}
#Levene's test for comparison of variances of exam scores in the two universities.
library(ggplot2)
library(car)
library(readr)
RExam <- read_csv("Data/RExam.csv")

rexam=RExam
head(rexam)

ggplot(rexam, aes(numeracy, fill=as.factor(uni)))+
  geom_histogram()+
  facet_wrap(~uni)



 leveneTest(rexam$lectures, rexam$uni)
leveneTest(rexam$lectures, rexam$uni, center = mean)

leveneTest(rexam$computer, rexam$uni)
leveneTest(rexam$numeracy, rexam$uni)




```




```{r transformation}

x=c(0,2,3,4,5,6,100,1000)
x
log(x+1)
sqrt(x)

head(log(dlf$day1))

head(dlf)
dlf$lgday1=log(dlf$day1)
head(dlf)
```



