---
title: "Ejerc_4b_geom_area_density"
author: "Raymond_L_Tremblay"
date: "2026-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```

## Cada Pregunta vale 5 puntos.  

------------------------------------------------------------------------

# 

# 1. Economic Example: Personal Savings Rate Over Time

Imagine you are a data analyst working for a small team within the **U.S. Bureau of Economic Analysis (BEA)**. Your group monitors long‑term financial trends to help economists and policy makers understand how Americans save money over time. One of the most important indicators your team studies is the **personal savings rate**—the percentage of personal income that individuals save rather than spend.

Your supervisor tells you:

> “We need a clear visual showing how the U.S. personal savings rate has changed over the years. Something intuitive, something that shows *flow* over time. Use one of the standard datasets we keep for training—look for the `economics` dataset inside the **ggplot2** package.”

You lookup the data and confirm that **`economics`** is a built‑in dataset that ships automatically with **ggplot2**. It contains monthly U.S. economic indicators collected from the Federal Reserve Bank of St. Louis (FRED), including:

-   `date` — monthly date values

-   `psavert` — *personal savings rate (%)*

-   `pce`, `unemploy`, `uempmed`, etc. (other variables you won’t need right now)

For your first task, your supervisor wants something simple but polished:

### **Your Mission**

Create a **filled area chart** showing how the personal savings rate (`psavert`) changes over time (`date`).

Specifically, you are expected to:

1.  Use **`geom_area()`** to draw the filled shape beneath the curve.

2.  Make sure the plot has clear axis labels and an informative title.

3.  Use a fill color with soft transparency so the shape feels like “economic volume” changing through time.

4.  Produce a figure that one could show to a policymaker and expect them to understand at a glance.

------------------------------------------------------------------------

## ✔️ Expected Output (description)

Students should produce a plot that shows:

-   **The x‑axis**: dates from the late 1960s to present

-   **The y‑axis**: % personal savings rate

-   **A filled area** under the curve

-   **A gentle color and light transparency**

*BONOS: add a horizontal dashed line at the median savings rate to provide a reference point for viewers.*

-   An overall impression of long‑term trends in U.S. saving behavior

The final result should look like a smooth “mountain range” of savings over time—simple, readable, and ready for interpretation.

```{r, eval=FALSE, echo=FALSE}
med <- median(economics$psavert, na.rm = TRUE)

ggplot(economics, aes(x = date, y = psavert)) +
  geom_area(fill = "#2C7FB8", alpha = 0.6, color = NA) +
  geom_hline(yintercept = med, linetype = "dashed", color = "gray40") +
  labs(title = "US Personal Savings Rate Over Time",
       y = "Personal savings rate (%)", x = NULL,
       caption = "Source: ggplot2::economics") +
  theme_minimal(base_size = 12)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# 2. Biological Example: Chick Weight Distributions by Diet

Imagine you are a young researcher in an animal physiology lab studying how diet influences growth in juvenile chickens. The lab ran a controlled experiment where newly hatched chicks were randomly assigned to four different diets. Each chick was weighed repeatedly as it grew. Your supervisor tells you:

“Before we analyze individual growth curves, let’s get a quick sense of the overall weight distribution for each diet. Use density curves — they're smooth, intuitive, and perfect for comparing population‑level differences.”

You open the R datasets and find the built‑in `ChickWeight` data frame in the `dataset` package. It contains:

### **Your Mission**

-   weight: body weight in grams

-   Time: age (days since birth)

-   Chick: individual ID

-   Diet — diet treatment (1, 2, 3, or 4)

### **Your Mission**

Since weight is a continuous variable and Diet is categorical, this is an ideal biological dataset for density plots. Your supervisor wants you to produce a visualization that:

Compares weight distributions between diets. Uses overlaid density curves, each with a different color. Uses transparency so overlapping regions remain visible. Helps determine whether certain diets tend to produce heavier or lighter chicks.

-   Generate a density plot where the area is **not filled**:

## ✔️ Expected Output (description)

Each diet has a different colored curve. - All curves are overlaid on the same axes.

\- X‑axis = chick weight

\- Y‑axis = density

\- Plot shows how the overall weight distribution differs across diets (e.g., Diet 3 tends to produce heavier chicks).

The final plot should visually communicate growth differences without requiring statistical modeling.

```{r, eval=FALSE, echo=FALSE}
library(ggplot2)

ggplot(ChickWeight, aes(x = weight, color = Diet)) +
  geom_density(linewidth = 1, alpha = 0.7) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Distribution of Chick Weights by Diet Treatment",
    x = "Chick weight (g)",
    y = "Density",
    color = "Diet"
  ) +
  theme_minimal(base_size = 12)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# 3. 🦷 Exercise — Density with Explicit Gaussian Kernel

**Dataset**: ToothGrowth **Goal**: Compare tooth length distributions across supplement types using geom_density(kernel = "gaussian") and discuss bandwidth effects.

## Story setup

You’re assisting in a nutrition/physiology lab testing whether vitamin C delivery method changes tooth growth in guinea pigs. Animals received vitamin C as either orange juice (OJ) or ascorbic acid (VC) at different doses. You want a quick, smooth comparison of tooth length (len) across supplement types to see if one tends to yield longer teeth.

### What students should produce that generates: DO ONE OF THESE ALTERNATIVES

## ✔️ Expected Output (description)

## Built-in biological dataset

data("ToothGrowth")

**Variables**:

\- len = tooth length (numeric, continuous)

\- supp = supplement type ("OJ" orange juice, "VC" ascorbic acid)

\- dose = dose level (0.5, 1, 2 mg)

-   An overlaid density plot of **len** by **supp**, using kernel = "gaussian".
-   A plot showing how bandwidth (via adjust) changes the smoothness/interpretation.

```{r, eval=FALSE, echo=FALSE}
library(ggplot2)

# Built-in biological dataset
data("ToothGrowth")
# Variables:
#   len  = tooth length (numeric, continuous)
#   supp = supplement type ("OJ" orange juice, "VC" ascorbic acid)
#   dose = dose level (0.5, 1, 2 mg)

# 1) Overlaid densities by supplement with explicit Gaussian kernel
ggplot(ToothGrowth, aes(x = len, color = supp)) +
  geom_density(kernel = "gaussian", linewidth = 1.1, alpha = 0.9) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Tooth Length by Supplement Type",
    subtitle = "Kernel density with explicit Gaussian kernel",
    x = "Tooth length (mm)",
    y = "Density",
    color = "Supplement"
  ) +
  theme_minimal(base_size = 12)

# 2) Show bandwidth sensitivity (narrower vs wider smoothing)
p_narrow <- ggplot(ToothGrowth, aes(len, color = supp)) +
  geom_density(kernel = "gaussian", adjust = 0.6, linewidth = 1.1) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Bandwidth: adjust = 0.6 (more wiggles)", x = "Tooth length", y = "Density") +
  theme_minimal(12)

p_wide <- ggplot(ToothGrowth, aes(len, color = supp)) +
  geom_density(kernel = "gaussian", adjust = 1.6, linewidth = 1.1) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "Bandwidth: adjust = 1.6 (smoother)", x = "Tooth length", y = "Density") +
  theme_minimal(12)

p_narrow
p_wide

# 3) Optional: Facet by dose to see how dose shifts distributions
ggplot(ToothGrowth, aes(len, color = supp)) +
  geom_density(kernel = "gaussian", linewidth = 1.0) +
  facet_wrap(~ dose, nrow = 1, labeller = label_both) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Tooth Length Distributions by Dose and Supplement",
    x = "Tooth length (mm)", y = "Density"
  ) +
  theme_minimal(12)
```
