---
title: "Ejerc_2_geom_point"
output:
  html_document:
    css: tutorial.css
    fig_caption: yes
    highlight: pygments
    theme: simplex
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```




This Assignment is to practice doing graphics using geom_point, geom_count and/or geom_histogram.

***

# Q1. 

Define what is the difference between geom_point, geom_count and geom_histogram.  In other words in what situation would you use each of these alternatives.

***

# Q2.
Use the data "Camas_Hospital" in the package **ggversa** and do a graphic with geom_point which relates the population size of countries and the number of hospital beds.

  a. The text of the axis should be in 15 pts. 
  b. The text of the axis should be in "bold".
  c. The text of the different axis "x" and "y" should be in different colours. 

```{r geom point-count, eval=FALSE, include=FALSE}
library(ggversa)
library(ggplot2)
#head(Camas_Hospital)



ggplot(Camas_Hospital,
       aes(Poblacion, Camas))+
  geom_point()+
  xlab("Poblacion")+
  ylab("Numero de Camas por 100,000")+
  theme(title=element_text(size=15,face="bold"))+
  theme(axis.title.x = element_text(colour="red"))+
  theme(axis.title.y = element_text(colour="blue"))

```


***

# Q3. 
Use the data "Camas_Hospital" in the package **ggversa** and do a graphic with geom_point which relates the population size of countries and the number of hospital beds.

  a. Ahora buscan la función para que el eje de x que sea el logaritmo de los valores en ese eje.
  
  No se olvide cambiar la información en el eje de "X" para que representa la información en log.

```{r geom point-count2, eval=FALSE, include=FALSE}
library(ggversa)
library(ggplot2)
#head(Camas_Hospital)



ggplot(Camas_Hospital,
       aes(log(Poblacion), Camas))+
  geom_point()+
  xlab("log(Poblacion)")+
  ylab("Numero de Camas por 100,000")+
  theme(title=element_text(size=15,face="bold"))+
  theme(axis.title.x = element_text(colour="red"))+
  theme(axis.title.y = element_text(colour="blue"))

```


