---
title: "Ejercicio de Practica"
output:
  html_document:
    css: tutorial.css
    fig_caption: yes
    highlight: pygments
    theme: simplex
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
# Primer asignación

### Este ejercicio es de practica.  No se evaluará el trabajo.
Se sugiere que Ud trata de cumplir con todos los pasos abajo.   El objetivo es de practicar las alternativas disponible en R, RSTUDIO y crear un documento .html.   

Es importante que Ud trata de hacer un documento profesional.   Esto tomara tiempo aprender y reconocer las diferentes funciones disponible en RStudio y RMarkdown!!!

Mi sugerencia es que trata de construir el documento y después habla con colegas para comparar como lo logró y resolver problemas.



#  Parte 1:

a. Instalar el package *ggversa*

 b. Mirar los datos **PIB_vs_Alfabetismo**, Producto Bruto Interno

```{r eval=FALSE, include=FALSE}
library(ggplot2)
library(ggversa)
head(PIB_vs_Alfabetismo)
```
 c. Hacer un gráfico que evalua la relación entre el Porciento_PIB (el prociento de producto interior dedicado a la educación) y el alfabetismo.  

```{r eval=FALSE, include=FALSE}
ggplot(PIB_vs_Alfabetismo, aes(Porciento_PIB, Alfabetismo))+
  geom_point()
```

4. Añade un color distincto a cada pais

```{r eval=FALSE, include=FALSE}
ggplot(PIB_vs_Alfabetismo, aes(Porciento_PIB, Alfabetismo, color=Pais))+
  geom_point()
```


# Parte #2

Algunos de estos ejercicio se vera miercoles. 

### Create a .csv files with the following information of your grandparent, uncles and siblings, have at least 10 people in the file (try to have many more)

### here is an example 
| Name    | Family_Name | Date_birth | Date_deceased | Number_children | Relationship |
|---------|-------------|---------------|-----------------|--------------------|--------------|
| Maria   | del_Valle   | 1882          | 1970            | 12                 | Grandmother  |
| Eugenio | del_Monte   | 1881          | 1956            | 13                 | Grandfather  |
|         |             |               |                 |                    |              |



  - Your Name - *in bold* (enegrecido)
  - Your Affiliation - **in italics**
  - Your Contact - email
  
    1. Import the file
    2. Show the link to the file in the chunk
    3. View the data
    4. Using the appropriate function show who was the person born first in your list
     head(df, n=1)
    5. Show the last 3 line of your data frame
    6. Calculate the mean life span of all your decease descendants
    7. Load a picture of your favorite grandparent and explain why this person had an impact or you found most interesting. 
    8. Make a numbered list of the class you enjoy the most to least.
    9. Include in your presentation the formula how to calculate the mean and the variance of a sample, using Latex formula.


```{r}
library(readr)
Practice_Data <- read_csv("Practice_Data.csv")
```


```{r}
head(Practice_Data)
```












